---
title: "NHANES Paired Chemical Exposure — EDA Report"
subtitle: "Datasets: PBCD_EXT, IHGEM, PFC, PFAS, SSPFAS, PHTHTE, UPHOPM · Cycles 2003–2017"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    theme: flatly
    highlight: tango
    self_contained: true
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.width = 10,
  fig.height = 5,
  dpi       = 150
)
```

```{r load-data, results='hide'}
# Load pre-built cache (run nhanes_paired_eda.R first to create it)
CACHE_FILE <- "nhanes_eda_cache.rds"
if (!file.exists(CACHE_FILE)) {
  stop(
    "Cache file not found: ", CACHE_FILE, "\n",
    "Please run nhanes_paired_eda.R first to download the data:\n",
    "  source('nhanes_paired_eda.R')"
  )
}

# Load cached data objects
cache         <- readRDS(CACHE_FILE)
merged_pairs  <- cache$merged_pairs
lod_summaries <- cache$lod_summaries
lab_store     <- cache$lab_store
needed        <- cache$needed

# Load plot helpers and EDA functions from the R script
# (source with local=TRUE so only functions/config are pulled in,
#  not the download sections which are guarded by SKIP_DOWNLOADS)
source("nhanes_paired_eda.R", local = TRUE)
```

---

# Overview

This report analyzes 10 dataset pairs from NHANES cycles 2003–2017. For each pair
**(A, B)**, participants measured in **both** datasets are identified and then examined:

- Distributions (raw and log₁₀-transformed) for chemicals in A and B
- Geometric means by cycle, sex, and age group
- Scatter plots and correlation heatmaps between A and B chemicals
- Limit-of-detection (LOD) summaries per chemical per cycle

## Pairs analyzed

```{r pair-table}
pair_tbl <- purrr::imap_dfr(eda_results, function(r, lbl) {
  data.frame(
    Pair        = lbl,
    Dataset_A   = r$a_key,
    Dataset_B   = r$b_key,
    N           = nrow(r$df),
    Cycles      = paste(sort(unique(r$df$year)), collapse = ", "),
    A_chemicals = length(r$cols_a),
    B_chemicals = length(r$cols_b)
  )
})
knitr::kable(pair_tbl, format.args = list(big.mark = ","))
```

---

```{r pair-sections, results='asis', fig.height=5}
# Loop over every pair and emit a full Markdown section dynamically
for (pair_label in names(eda_results)) {
  r     <- eda_results[[pair_label]]
  df    <- r$df
  a_key <- r$a_key
  b_key <- r$b_key

  cat("\n\n---\n\n")
  cat(sprintf("# %s\n\n", pair_label))
  cat(sprintf(
    "**Dataset A:** %s &nbsp; | &nbsp; **Dataset B:** %s &nbsp; | &nbsp; **N (overlap):** %s &nbsp; | &nbsp; **Cycles:** %s\n\n",
    a_key, b_key, format(nrow(df), big.mark = ","),
    paste(sort(unique(df$year)), collapse = ", ")
  ))

  # ── Demographics snapshot ──────────────────────────────────────────────────
  cat("## Demographics\n\n")

  demo_tbl <- df %>%
    summarise(
      N           = n(),
      `Pct Male`  = round(mean(gender == "Male",   na.rm = TRUE) * 100, 1),
      `Pct Female`= round(mean(gender == "Female", na.rm = TRUE) * 100, 1),
      `Median Age`= round(median(RIDAGEYR, na.rm = TRUE), 1),
      `Pct <18`   = round(mean(RIDAGEYR < 18,  na.rm = TRUE) * 100, 1),
      `Pct 18-44` = round(mean(RIDAGEYR >= 18 & RIDAGEYR < 45, na.rm = TRUE) * 100, 1),
      `Pct 45-64` = round(mean(RIDAGEYR >= 45 & RIDAGEYR < 65, na.rm = TRUE) * 100, 1),
      `Pct 65+`   = round(mean(RIDAGEYR >= 65, na.rm = TRUE) * 100, 1)
    )
  print(knitr::kable(demo_tbl, format.args = list(big.mark = ",")))
  cat("\n\n")

  # ── LOD summaries ──────────────────────────────────────────────────────────
  cat("## Limit of Detection\n\n")
  cat("### Dataset A:", a_key, "\n\n")
  lod_a <- r$lod_a
  if (!is.null(lod_a) && nrow(lod_a) > 0) {
    print(knitr::kable(lod_a, format.args = list(big.mark = ",")))
    p <- plot_lod(lod_a, a_key)
    if (!is.null(p)) print(p)
  } else { cat("_(No LOD flag columns found)_\n\n") }

  cat("\n\n### Dataset B:", b_key, "\n\n")
  lod_b <- r$lod_b
  if (!is.null(lod_b) && nrow(lod_b) > 0) {
    print(knitr::kable(lod_b, format.args = list(big.mark = ",")))
    p <- plot_lod(lod_b, b_key)
    if (!is.null(p)) print(p)
  } else { cat("_(No LOD flag columns found)_\n\n") }

  # ── Distributions for Dataset A ────────────────────────────────────────────
  cat("\n\n## Distributions — Dataset A:", a_key, "\n\n")
  for (col in r$cols_a) {
    cat("###", col, "\n\n")
    vals <- df[[col]][!is.na(df[[col]]) & df[[col]] > 0]
    if (length(vals) < 5) { cat("_(insufficient data)_\n\n"); next }

    # Summary stats
    stats_tbl <- data.frame(
      N        = length(vals),
      Min      = round(min(vals), 4),
      P25      = round(quantile(vals, 0.25), 4),
      Median   = round(median(vals), 4),
      P75      = round(quantile(vals, 0.75), 4),
      Max      = round(max(vals), 4),
      GeoMean  = round(geo_mean(vals), 4),
      `Pct>0`  = round(mean(vals > 0) * 100, 1)
    )
    print(knitr::kable(stats_tbl, format.args = list(big.mark = ",")))
    cat("\n\n")
    plot_dist(df, col, title_prefix = paste0(a_key, " · "))
    cat("\n\n")
    p <- plot_gm_by_cycle(df, col, a_key)
    if (!is.null(p)) print(p)
    cat("\n\n")
    p <- plot_gm_by_demo(df, col, a_key)
    if (!is.null(p)) print(p)
    cat("\n\n")
  }

  # ── Distributions for Dataset B ────────────────────────────────────────────
  cat("\n\n## Distributions — Dataset B:", b_key, "\n\n")
  for (col in r$cols_b) {
    cat("###", col, "\n\n")
    vals <- df[[col]][!is.na(df[[col]]) & df[[col]] > 0]
    if (length(vals) < 5) { cat("_(insufficient data)_\n\n"); next }

    stats_tbl <- data.frame(
      N        = length(vals),
      Min      = round(min(vals), 4),
      P25      = round(quantile(vals, 0.25), 4),
      Median   = round(median(vals), 4),
      P75      = round(quantile(vals, 0.75), 4),
      Max      = round(max(vals), 4),
      GeoMean  = round(geo_mean(vals), 4),
      `Pct>0`  = round(mean(vals > 0) * 100, 1)
    )
    print(knitr::kable(stats_tbl, format.args = list(big.mark = ",")))
    cat("\n\n")
    plot_dist(df, col, title_prefix = paste0(b_key, " · "))
    cat("\n\n")
    p <- plot_gm_by_cycle(df, col, b_key)
    if (!is.null(p)) print(p)
    cat("\n\n")
    p <- plot_gm_by_demo(df, col, b_key)
    if (!is.null(p)) print(p)
    cat("\n\n")
  }

  # ── A vs B scatter plots ───────────────────────────────────────────────────
  if (length(r$cols_a) > 0 && length(r$cols_b) > 0) {
    cat("\n\n## A vs B Scatter Plots\n\n")
    # Plot up to the top 3 A-chemicals × top 3 B-chemicals
    top_a <- head(r$cols_a, 3)
    top_b <- head(r$cols_b, 3)
    for (ca in top_a) {
      for (cb in top_b) {
        p <- plot_scatter(df, ca, cb)
        if (!is.null(p)) { print(p); cat("\n\n") }
      }
    }
  }

  # ── Correlation heatmap ────────────────────────────────────────────────────
  if (length(r$cols_a) + length(r$cols_b) >= 2) {
    cat("\n\n## Correlation Heatmap (log-transformed)\n\n")
    cat("Each cell shows Pearson correlation after log1p transform. Upper triangle shown.\n\n")
    plot_correlation(df, r$cols_a, r$cols_b, pair_label)
    cat("\n\n")
  }
}
```

---

# Session Info

```{r session}
sessionInfo()
```
